Search for Spatially Extended Sources Using Two Years of Flight Data

### 2011-04-12 extended_catalog/paper/extended_catalog.tex
Working on paper

\documentclass[12pt]{article}
\usepackage{fullpage}

@@

One can define a parameter $\text{TS}_\text{ext}$ as twice the increase
in log likelihood going from fitting a source with a point hypothesis to fitting
it as an extended hypothesis. So $\text{TS}_\text{ext}=2\times(LL_\text{ext}-LL_\text{ps})$.
Note that Wilk's Theorem only holds when the likelihood is maximized simultaneoulsy
fitting all of the parameters of model. It is for this reason why {\em pointlike}
goes through such great effort to simultaneously fit the position and extension of
a source.
In going from fitting a point source to fitting a simple extended source such as
a disk, one degree of freedom, the source width, is added. According to
Wilk's Theorem, the distribution of $\text{TS}_\text{ext}$ for no signal (a true
point source) should follow a $\chi^2$ distribution with one degree of freedom\cite{Wilks_Theorem}.
On the other hand, since the null hypothesis exists on the edge of parameter space,
and so so Wilk's Theorem does not hold\cite{Warnings about Wilk's Theorem Paper}.
One might expect the actual distribution to be a half $\chi^2$ distribution with a delta
function at 0. This was the distribution found to be true for source finding of
gamma ray sources.\cite{Mattox_et_All_Paper}. A Monte Carlo study has been performed to
show how well we can believe the statistical significance.

@@

\includegraphics{k

>>>>

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

@@

To calculate the statistical significance of the extension of a source,
one can define a parameter $\text{TS}_\text{ext}$ as twice the increase
in log likelihood between fitting a source with a
spatially extended hypothesis and a point hypothesis. 
So $\text{TS}_\text{ext}=2\times(LL_\text{ext}-LL_\text{ps})$.

In going from fitting a point source to fitting a simple extended source such as
a disk, the source's width is the only degree of freedom added to the model. According to
Wilk's Theorem, the distribution of $\text{TS}_\text{ext}$ for no signal (when there is a true
point source) should follow a $\chi^2$ distribution with one degree of freedom\cite{Wilks_Theorem}.
Wilk's Theorem only holds when the likelihood is maximized simultaneoulsy
fitting all of the parameters of model. It is for this reason why {\em pointlike}
goes through such great effort to simultaneously fit the position, extension,
and spectral paramters of a source.

On the other hand, since the null hypothesis rests on the edge of
parameter space, when the extension is fixed to 0, Wilk's Theorem does
not formally hold\cite{Warnings about Wilk's Theorem Paper}.  One might
expect the actual distribution to be a half $\chi^2_1$ distribution with
one degree of freedom plus half of a delta function at 0:
\begin{equation}
  P(\text{TS})=\tfrac{1}{2}(\chi^2_1(\text{TS})+\delta(\text{TS})).
\end{equation}
The false detection probability can be estimated by integrating this function 
from an observed test statistic value to infinity. It is for this
particular distribution that the often quoted result holds that
$\sqrt{TS}$ is a measure of the number of $\sigma$ of the detection.

The half $\chi^2_1$ distribution was found by a monte carlo study
to hold for the special case of source finding of EGRET sources using the
maximum likelihood technique.  They were comparing the hypothesis of a
source to the hypothesis of just a background\cite{Mattox_et_All_Paper}.
Their argument for why this null hypothesis distribution was reasonable
was that half of the time, statistical underfluctuations would cause
there to be fewer fewer photons then expected where the sources is being
tested. Since the maximim likelihood machinery cannot fit negative
sources (which would lead to negative model predicted for which the
poisson likelihood is undefined), half of the tests would not have the
likelihood improved by adding a source to the model and $\text{TS}=0$.
It is for the case that the null distribution has

It seems plausable that a similar situation could happen for when 
testing a source for extension. Half of the time, statistical flucutations
might make the observed counts narrower thant the point spread function
and the likelhood would not improve by fitting the source with an extended
hypothesis. We tested this hypothesis using a dedicated monte carlo study.

\begin{figure}
\includegraphics{mc_plots/ts_ext_emin_1000.pdf}
% this plot came from /u/gl/lande/work/fermi/extended_catalog/monte_carlo/ts_ext/v3/plot.py
% using data from /nfs/slac/g/ki/ki03/lande/extended_catalog/monte_carlo/ts_ext/v3_emin_1000/merged.hdf5
\caption{Lalla}
\end{figure}

@@

\section{Extended Source Sensitivity}
